Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Signal

### 2020-08-06 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
Advances

\title{Robust Digital Envelope Estimation Via Geometric Properties of a Arbitrary Signal}
\author{
  Carlos Tarjano
  \and 
  Valdecy Pereira
  }

@@

  % background  
  Despite being an elusive concept, mathematically well defined only for artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
  intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape. It is implied in this definition that the envelope is non periodic in general and is thus better addressed by specific techniques diverse from those used in the analysis of the periodic, rapidly changing wave it encompasses.
  % motivation
  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design, based on prior knowledge about the wave under investigation.
  % objective
  In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
  % methods
  The approach here described draws inspiration from general and differential geometry to isolate the frontier of an arbitrary signal. The concepts of alpha-shapes, concave hulls, arc-length parametrization and discrete curve curvature are adapted. We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and subsequently, the complexity of the algorithm.
  % results

@@

\section{Introduction} % 500 words
% Opener sentence
Despite being ubiquitous in digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications.
% Literature review | Brief Context of Prior Research
In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones. That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning. The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.

% enumerate some uses in diverse areas

Also, the envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a subsequent spectral analysis. In the context of sound synthesis, for example, one would be able to apply different methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and an different approach to the periodic and relatively fast changes of the temporal fine structure.

% Cite current algorithms

% Restate Your Question as Something Not Known or Fully Understood by Prior Research
Thus, in this paper we formulate a general approach to envelope detection, examining the intrinsic characteristics of a generic, spectrally complex wave in order to avoid parameter tuning.
% State the Significance of Your Question
While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.
% State Your Claim | Objective | Hypothesis
% brief outlook on the structure of the paper

@@

As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent. this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units.

@@

For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $c(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $. This definition provides a welcome link between algebra and geometry, but also poses two difficulties to our particular case that need to be addressed: the fact that our signal isn't even continuous, and thus neither smooth, and that we don't have a solid unit for the amplitude.

\subsection{Theory}

>>>>

\title{Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Signal}
\author{
  Carlos Tarjano
  \and 
  Valdecy Pereira
  }

@@

  % background  
  Despite being an elusive concept, mathematically well defined only for artificially generated waves, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information carrying medium in spoken voice and telecommunications, for example.
  intuitively, the temporal envelope can be understood as a smooth function on the same variable as the principal wave, or temporal fine structure, that modulates the signal, responsible for its outer shape. It is implied in this definition that the envelope is non periodic in general and is thus better addressed by specific techniques diverse from those used in the analysis of the periodic, rapidly changing wave it encompasses.
  % motivation
  Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. Despite of that, a general approach to envelope detection of signals with rich spectral content seems to be lacking, and most methods involve manual intervention, like filter design, based on prior knowledge about the wave under investigation.
  % objective
  In this paper we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
  % methods
  The approach here described draws inspiration from general, computational and differential geometry to isolate the frontier of an arbitrary signal. The concepts of alpha-shapes, concave hulls, arc-length parametrization and discrete curve curvature are adapted. We also define a pulse in the context of a arbitrary digital signal as a means to reduce dimensionality and the complexity of the proposed algorithm.
  % results

@@

\section{Introduction} % 500 words
% Opener sentence
Despite being ubiquitous in digital signal processing, the literature about envelope detection is very fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise contents, in line with the most common usages of those algorithms with artificial signals in the context of analog telecommunications.
% Literature review | Brief Context of Prior Research
In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: According to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of the soundwave in the context of the intelligibility of mandarin tones. That is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning. The envelope helps to convey emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics of concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.
% Cite current algorithms
When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.

If one adresses the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection in the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $. For that purpose \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well defined extension to the convex hull of a finite set of points, related to the Delaunay triangulation and Voronoi diagrams of those points.

This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, that involves the estimation and removal of the Blaze function -a kind of envelope- of a echelle spectrograph, being particularly align with what is intended here.

% Restate Your Question as Something Not Known or Fully Understood by Prior Research
Thus, in this paper we formulate a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex, wave in order to avoid the need to manual intervention or parameter tuning.
% State the Significance of Your Question
While the robustness of the proposed approach allows it to be used as a plug in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis. 
The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and a accurate description of the envelope would be useful for a cleaner spectral analysis. Moreover, the algorithm naturally divides a signal in pseudo-cycles, that could serve pragmatic building blocks for the reconstruction of the fine structure of the wave.
% State Your Claim | Objective | Hypothesis
In the context of sound synthesis, one would be able to apply specific methods for the recreation of the envelope, more in line with its smooth, slow varying and non periodic nature and a different approach to the periodic and relatively fast changes of the temporal fine structure.
% brief outlook on the structure of the paper

@@

% pulse, simplified pulse, pseudo-cycle image

As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent. this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units. One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.

@@

Besides, those definitions will allow us to derive the frontier of a wave using an alpha-shapes inspired algorithm without the need to compute the Delaunay triangulation first. Translating the intuitive explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree}, we can say that the points in the frontier are those touched by a circle outside the signal that is not allowed to contain any point of the signal; one can picture a circle being rolled above (or below, in the case of the negative envelope) the signal, and marking the points it touches as frontier points.

% snowball image

It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the instantaneous curvature of a discrete function is needed. Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods, or the calculation of the curvature of a fitted curve \parencite{2001CoeurjollyDiscrete}. As we are interested in the average curvature, the second approach is more natural, since it enables us to calculate the average curvature via the integral of the curvature in the fitted curve.

% osculating circle image, fitted line

For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $. Since we are assuming smoothness of envelope in the neighbourhood of a pulse, it makes sense to approximate it with a parabola $ y(x)=ax^2+bx+c $, since a higher order polynomial would provide unnecessary complexity. Besides some characteristics of the parabola derived from its relation with Bezier curves will be useful later.

Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.

\begin{equation} \label{eq:1}
  \overline{\kappa}(x) = 
  \frac{
      \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
  }{
    (x_1 - x_0)
  }
\end{equation}

% \subsection{Theory}
